Statement of Intent

### 2010-12-11 applications/phd_programs/nsf_research.tex
Started working on my phd essay

This project started when a physics professor approached my advisor seeking
help solving a crystallography problem.  As a new student in the lab, I
accepted the problem mostly in an auxiliary role, writing some code and running
experiments for Dr. Campbell, the physics professor.  We used particle swarm
optimization (PSO), a bio-inspired optimization algorithm similar to
evolutionary algorithms, to try to fit to data his model of what was happening
in a particular kind of crystal.  After spending a lot of time seeing the
output of the algorithm, I also helped to interpret the results.  I discovered
that we had too many free parameters in our model and helped Dr.  Campbell to
make progress in solving his problem.  Because of our help in solving his
problem, Dr. Campbell has made what he considers a breakthrough in
understanding these crystals and plans to submit his work for publication soon.

The crystallography problem we were solving required the evaluation of a
function that took several minutes to run.  As each particle in the algorithm
must evaluate the function at every iteration, we used several hundred
processors on BYU's supercomputer in our optimization of the function.  This
led us to ask some interesting questions about how we could best use those
processors to solve the problem.  We first experimented with the topology of
the algorithm, or how the particles in the swarm communicate with each other.
I did much of the experimentation, though I was directed by a graduate student
in the lab, who eventually wrote the paper that we
published~\cite{mcnabb-2009-large-particle-swarms}.  We saw some improvements
in the performance of the algorithm with our new topologies, though we also
noticed that increasing the number of particles with additional processors had
diminishing returns.

@@

\footnotesize
\bibliographystyle{plain}
\renewcommand\bibsection{\noindent \small\textbf{Publications}\vspace{-2mm}\footnotesize}
\bibliography{../../../bib/aml/bib}

>>>>

This project started when Dr. Campbell, a physics professor,  approached my
advisor seeking help solving a crystallography problem.  As a new student in
the lab, I accepted the problem mostly in an auxiliary role, writing some code
and running experiments for Dr. Campbell.  We used particle swarm optimization
(PSO), a bio-inspired optimization algorithm similar to evolutionary
algorithms, to try to fit to data his model of what was happening in a
particular kind of crystal.  After spending a lot of time seeing the output of
the algorithm, I also helped to interpret the results.  I discovered that we
had too many free parameters in our model and helped Dr.  Campbell to make
progress in solving his problem.  Because of our help, Dr.  Campbell has made
what he considers a breakthrough in understanding these crystals and plans to
submit his work for publication soon.

The crystallography problem we were solving required the evaluation of a
function that took several minutes to run.  As each particle in the algorithm
must evaluate the function at every iteration, we used several hundred
processors on BYU's supercomputer in our optimization of the function.  This
led us to ask some interesting questions about how we could best use those
processors to solve the problem.  We first experimented with the topology of
the algorithm, or how the particles in PSO communicate with each other.  I did
much of the experimentation, though I was directed by a graduate student in the
lab, who eventually wrote the paper that we
published~\cite{mcnabb-2009-large-particle-swarms}.  We saw some improvements
in the performance of the algorithm with our new topologies, though we also
noticed that increasing the number of particles with additional processors gave
diminishing returns.

@@

\footnotesize
\bibliographystyle{plain}
\renewcommand\bibsection{\noindent \small\textbf{Publications}\vspace{-2mm}\footnotesize}
\bibliography{bib}
